Paper

### 2025-03-12 paper/paper.qmd
discussion

The model was designed to simulate the spread of infection/colonisation at a baseline (or background rate), followed by an outbreak period of increased transmission, in turn followed by a period of enhanced infection control to return back to the baseline rate. The model was calibrated to a report of a _C. difficile_ outbreak reported by @salgadoAnalysisOutbreakClostridium2009. Approximate Bayesian computation was used to calibrate the model to the reported rates of infection. When model inputs were drawn from the posterior distributions, the model was able to simulate plausible outbreak scenarios that were similar to the observed rates. A sensitivity analysis was performed over parameters that represent infection control interventions (such as the frequency of toilet cleaning), demonstrating the possible applications of this model for informing operational infection control practices. Interventions that controlled the spread of infection, such as the cohorting of patients to single-occupancy beds, and the increased frequency of toilet cleaning, led to a reduced number of cases during the outbreak period.

Repeated simulations using the model generated nosocomial CDI rates that showed a clear outbreak peak, followed by control of colonisation/infection through a change in infection control practice, resembling a real hospital CDI outbreak. The interactive nature of the model can be used for outbreak preparedness by testing out different infection control strategies. In summary, we report a pathogen-agnostic model that can be used to simulate any infection with features such as faecal-oral transmission, colonisation, and antimicrobial selective pressure effects.

>>>>

The model was designed to simulate the spread of infection/colonisation at a baseline (or background) rate, followed by an outbreak period of increased transmission, in turn followed by a period of enhanced infection control to return back to the baseline rate. The model was calibrated to a report of a _C. difficile_ outbreak reported by @salgadoAnalysisOutbreakClostridium2009. Approximate Bayesian computation was used to calibrate the model to the rates of infection reported in the paper. When model inputs were drawn from the posterior distributions, the model was able to simulate plausible outbreak scenarios that were similar to the observed rates. A sensitivity analysis was performed over parameters that represent infection control interventions (such as the frequency of toilet cleaning), demonstrating the possible applications of this model for informing operational infection control practices. Interventions that controlled the spread of infection, such as the cohorting of patients to single-occupancy beds, and the increased frequency of toilet cleaning, led to a reduced number of cases during the outbreak period.

To simulate outbreaks and periods of enhanced infection control, the model was designed to change parameters at discrete time points. For example, the rate of antimicrobial prescriptions was higher during the outbreak period, promoting the spread of infection and colonisation. Later, the prescription rate was reduced to help control the outbreak. The NetLogo model implementation was designed so that a fixed parameter set was used as an input, and the outbreak and enhanced IPC parameters were only used during their respective periods. In effect, all input parameters were provided as a fixed input -- some parameters were only used by the simulation at defined discrete intervals. This approach may be unsuitable for modelling of transmission without a clear outbreak pattern -- more complex implementations of time-varying parameters could be explored.

Although agent-based models have been shown to be useful for simulating and modelling infectious disease transmission, they have limitations. The model described here is a simplification of what is likely to be a highly intricate and complicated system. For example, transmission was assumed to be 'random' or acquired through the use of a contaminated toilet. In reality, other transmission pathways likely exist, such as through contact with staff or contaminated bedding. Antimicrobial exposure was considered to be a binary risk factor for an agent becoming infected or colonised. Antimicrobial class and duration of exposure are factors that are not accounted for in this model. An agent-based model that captures all the details that influence the spread of infection in a hospital setting is unlikely to be feasible, due to the computational challenge and the large number of parameters that would need to be estimated. A reasonable approach is for the model to be as simple as possible, while still capturing the features required to answer the research question [@sunSimpleComplicatedAgentbased2016]. Another limitation of agent-based models is the requirement for detailed data to calibrate and validate the model. Here, the data used was primarily from a retrospective descriptive dataset. As such, information on certain parameters was not available (e.g., the proportion of patients that were administered an antimicrobial during their admission) -- these parameters were estimated from the literature. An important future direction would be to prospectively inform and validate the model with data collected specifically for this purpose.

Agent-based models that simulate the hospital transmission of pathogens transmitted through the faecal-oral route, including _C. difficile_ and vancomycin-resistant _Enterococci_, have been reported previously [@codellaAgentbasedSimulationModel2015; @debosckerAgentbasedModelSimulate2021]. Some models have also explored the impact of one or more infection control interventions on the spread of infection [@barkerInterventionsReduceIncidence2018; @rubinSimulationBasedAssessmentStrategies2013]. Our model builds on these works by designing and calibrating the model to an outbreak and enhanced IPC period. The motivation for this was to design a model that can be used to support infection and prevention teams in the following: 1) preparedness for outbreak situations; and 2) decision-making on the implementation of infection control interventions during an outbreak. Furthermore, the model was designed to be pathogen-agnostic to allow for a stepwise approach for future work -- pathogen-specific features (e.g., diagnostic and laboratory testing) can be added as required. The choice of NetLogo as a platform allows for the model to be used interactively by non-programmers, such as infection control practitioners; the model can also be modified and extended with the relatively simple NetLogo language.

In conclusion, the agent-based model described here can simulate outbreaks of infection in a hospital setting, and estimate the impact of different infection control interventions on the number of nosocomial infections. Although the model was calibrated to a _C. difficile_ outbreak, it can be adapted for use in other pathogens that are transmitted through the faecal-oral route under antimicrobial selective pressure, such as vancomycin-resistant _Enterococci_ and extended-spectrum beta-lactamase-producing Enterobacterales.
